---
title: "Report"
date: "12/5/2018"
output: github_document
---

# Data

## Data Source

Our [data](https://www.kaggle.com/osmi/mental-health-in-tech-2016/downloads/mental-health-in-tech-2016.zip/1) comes from [OSMI Mental Health in Tech Survey](https://osmihelp.org/research/) which we first find in [kaggle](https://www.kaggle.com/osmi/mental-health-in-tech-2016).

This data includes 1433 observations and 63 variables.

## Data cleaning

```{r}
library(tidyverse)
```

First, we should import data.

```{r}
mentalhealth_2016_df = read_csv("./data/mental_health_in_tech_2016.csv") 
```

Variables name in this data set are questions, so we change it to [short words](Variable name.pdf) for easy-reading.

```{r}
colnames(mentalhealth_2016_df) <- c("self_employed", "num_employees", "tech_company", "tech_role", "benefits", "care_options", "employer_discussion", "employer_help", "anonymity", "medical_leave", "mental_health_consequences", "physical_health_consequences", "coworkers_discussion", "supervisor_discussion", "mental_vs_physical", "obs_consequence", "medical_coverage", "help_resourcces", "whether_reveal_business_contacts", "reveal_concequences_business_contects", "whether_reveal_coworkers", "reveal_concequences_coworkers", "productivity_affect", "work_time_affected", "preemployers", "preemployers_benefits", "preemployers_care_options", "preemployers_discussion", "preemployer_help", "pre_anonymity", "pre_mental_health_consequences", "pre_physical_health_consequences", "pre_coworkers_discussion", "pre_supervisors_discussion", "pre_mental_vs_physical", "pre_obs_consequence", "physical_health_interview", "physical_health_interview_reason", "mental_health_interview", "mental_health_interview_reason", "career_influence", "coworkers_view", "friends_family_share", "unsupportive_badly_handled", "less_likely_reveal", "family_history", "mental_health_previous", "mental_health_now", "condition_diagnosed", "possible_condition", "professional_diagnosed", "condition_professional_diagnosed", "seek_treatment", "work_interferes_treated", "work_interferes_untreated", "age", "gender", "country_live", "territory_live", "country_work", "territory_work", "work_position_kind", "work_remotely")
```

The age of one observation is 323 which is impossible, so we exclude this observation. Besides, there are lots of different forms of answers in `gender` variable which makes analysis process more difficult, so we use 'Female' and 'Male' to represent gender in each observation.

```{r}
mentalhealth_2016_tidied =
  mentalhealth_2016_df %>% 
  filter(age < 100) %>% 
  mutate(gender = str_replace(gender, "^[Ww]oman$", "Female"),     # female
         gender = str_replace(gender, "^[Ff]$", "Female"),
         gender = str_replace(gender, "^fem$", "Female"),
         gender = str_replace(gender, "^female$", "Female"),
         gender = str_replace(gender, "^female/woman$", "Female"),
         gender = str_replace(gender, "^fm$", "Female"),
         gender = str_replace(gender, "AFAB", "Female"),
         #gender = str_replace(gender, "AFAB", "Female"),
         gender = str_replace(gender, "^[Mm]$", "Male"),  # male
         gender = str_replace(gender, "^mail$", "Male"),  #?
         gender = str_replace(gender, "^MALE$", "Male"),
         gender = str_replace(gender, "^Male.$", "Male"),
         gender = str_replace(gender, "^Malr$", "Male"),    #?
         gender = str_replace(gender, "^[Mm]an$", "Male"),
         #gender = str_replace(gender, "^man$", "Male"),
         gender = str_replace(gender, "^male$", "Male"))

head(mentalhealth_2016_df)
```

